---
output:
  html_document:
    theme: spacelab
  pdf_document:
    keep_tex: no
  word_document: default
---
Class summary
========================================================
```{r echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=4, fig.align='center')
options(digits=2)
source("helpers.R")
```

```{r}
mid.1 <- makeTest("catalyst3.csv", "Midterm 1", 48)
mid.2 <- makeTest("catalyst3.csv", "Midterm 2", 40)
final <- makeTest("catalyst3.csv", "Final", 100)
```

## Midterm 1

There were a total of `r mid.1$total.points` points on the exam.

The median is `r median(mid.1$curr.test$total)`, the mean is `r mean(mid.1$curr.test$total)`, and the standard deviation is `r sd(mid.1$curr.test$total)`. Here are a few plots showing distributions of points on each question.

Here's a box plot for each question.
```{r}
print(questionBoxPlot(mid.1))
```

Here are histograms for each question (down the diagonal), scatter plots for pairs of questions (under the diagonal), and the correlation coefficients for each pair of questions (above the diagonal). You can understand a pair of questions by looking in one row and one column. Let us know if you can't figure out what this means.
```{r}
print(pairPlot(mid.1))
```

Here's a histogram of the overall test performance.
```{r}
print(testHistogram(mid.1))
```

This last plot compares the probability density derived from class performance with a normal distribution having the same mean and standard deviation; it is a smoother representation of how the class distribution compares to a normal.

```{r}
print(testDensityPlot(mid.1))
```

## Midterm 2

There were a total of `r mid.2$total.points` points on the exam.

The median is `r median(mid.2$curr.test$total)`, the mean is `r mean(mid.2$curr.test$total)`, and the standard deviation is `r sd(mid.2$curr.test$total)`. Here are a few plots showing distributions of points on each question.

Here's a box plot for each question.
```{r}
print(questionBoxPlot(mid.2))
```

Here are histograms for each question (down the diagonal), scatter plots for pairs of questions (under the diagonal), and the correlation coefficients for each pair of questions (above the diagonal). You can understand a pair of questions by looking in one row and one column. Let us know if you can't figure out what this means.
```{r}
print(pairPlot(mid.2))
```

Here's a histogram of the overall test performance.
```{r}
print(testHistogram(mid.2))
```

This last plot compares the probability density derived from class performance with a normal distribution having the same mean and standard deviation; it is a smoother representation of how the class distribution compares to a normal.

```{r}
print(testDensityPlot(mid.2))
```

## Final

There were a total of `r final$total.points` points on the exam.

The median is `r median(final$curr.test$total)`, the mean is `r mean(final$curr.test$total)`, and the standard deviation is `r sd(final$curr.test$total)`. Here are a few plots showing distributions of points on each question.

Here's a box plot for each question.
```{r}
print(questionBoxPlot(final))
```

Here are histograms for each question (down the diagonal), scatter plots for pairs of questions (under the diagonal), and the correlation coefficients for each pair of questions (above the diagonal). You can understand a pair of questions by looking in one row and one column. Let us know if you can't figure out what this means.
```{r}
print(pairPlot(final))
```

Here's a histogram of the overall test performance.
```{r}
print(testHistogram(final))
```

This last plot compares the probability density derived from class performance with a normal distribution having the same mean and standard deviation; it is a smoother representation of how the class distribution compares to a normal.

```{r}
print(testDensityPlot(final))
```


# Comparisons

Here's a scatter plot of the total scores on the two midterms.
```{r}
print(compareTwoTests(mid.1, mid.2))
```

Here are plots comparing the midterms and final.
```{r}
print(compareTwoTests(mid.1, final))
print(compareTwoTests(mid.2, final))
```

