---
output:
  html_document:
    theme: spacelab
  pdf_document:
    keep_tex: no
  word_document: default
---
Class summary
========================================================
```{r echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=4, fig.align='center')
```

```{r}
library(xlsx)
library(reshape)
library(ggplot2)

dehex <- function(x) {
  switch(x, 
         "0" = 0,
         "1" = 1,
         "2" = 2,
         "3" = 3,
         "4" = 4,
         "5" = 5,
         "6" = 6,
         "7" = 7,
         "8" = 8,
         "9" = 9,
         a = 10,
         b = 11,
         c = 12,
         d = 13,
         e = 14,
         f = 15)
}

grades <- function (x, questions) {
    if (x == "" | is.na(x)) {
      rep(NA, times = questions)
    } else {
    L <- sapply(strsplit(tolower(x), split = "")[[1]], dehex)
  names(L) <- paste("Q", 1:questions, sep=".")
  L
  }
}

# need to deal with NA right above and below
test <- function (x) {
	questions <- 4
	data.frame(t(sapply(x, function(y) grades(y, questions))))
}

makeTest <- function (q) {
  function(x) test(x, q)
}

# data ingestion
# START ROW DEPENDS ON CATALYST
# END ROW DEPENDS ON NUMBER OF STUDENTS
# WEBASSIGN EXPORT IS ESSENTIALLY BROKEN, SO HAVE TO CONVERT TO CSV
# THEN THOSE MORONS ENCODED A COLUMN OF PERCENTAGES USING %!!!!
cats <- read.xlsx("catalyst2.xls", 1, colClasses="character", startRow=3)[-c(1,2),]
cats <- cats[1:117,]
```

Here's a summary of how the class did on the midterm.

## Midterm 2

```{r}
mid.1 <- test(cats$Midterm.2)
mid.1 <- mid.1[complete.cases(mid.1),]
qs <- ncol(mid.1)
mid.1$total <- rowSums(mid.1)
curr.test <- mid.1
mid.1.points <- 40
# exclude the 0s; this probably means something was very wrong, like students absent
curr.test <- curr.test[curr.test$total != 0,]
```

There were a total of `r mid.1.points` points on the exam.

The median is `r median(curr.test$total)`, the mean is `r mean(curr.test$total)`, and the standard deviation is `r sd(curr.test$total)`. Here are a few plots showing distributions of points on each question. Keep in mind that the second two problems were each out of 8 points, so the box plots, etc., should be lower. (Perhaps in a future revision I will normalize to show relative performance, regardless of point value.)


```{r}
ggplot(melt(curr.test[,1:qs])) + geom_boxplot(aes(x=variable, y=value)) + xlab("Problem") + ylab("Points")
```
```{r}
ggplot(melt(curr.test[,1:qs])) + geom_histogram(aes(x=value), binwidth=1) + facet_grid(.~variable)
```

```{r}
ggplot(curr.test) + geom_histogram(aes(x=total), binwidth=2)
```

This last plot compares the probability density derived from class performance with a normal distribution having the same mean and standard deviation; it is a smoother representation of how the class distribution compares to a normal.

```{r}
df <- curr.test[curr.test$total != 0, ]
ggplot(df, aes(x=total)) + geom_density() + 
  stat_function(fun=dnorm, args=list(mean=mean(df$total), sd=sd(df$total)), color='red')
```