Class summary
========================================================
```{r echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=4, fig.align='center')
```

```{r}
library(xlsx)
library(reshape)
library(ggplot2)

dehex <- function(x) {
  switch(x, 
         "0" = 0,
         "1" = 1,
         "2" = 2,
         "3" = 3,
         "4" = 4,
         "5" = 5,
         "6" = 6,
         "7" = 7,
         "8" = 8,
         "9" = 9,
         a = 10,
         b = 11,
         c = 12,
         d = 13,
         e = 14,
         f = 15,
         g = 16,
         h = 17, 
         i = 18,
         j = 19,
         k = 20)
}

grades <- function (x, questions) {
    if (x == "" | is.na(x)) {
      rep(NA, times = questions)
    } else {
    L <- sapply(strsplit(tolower(x), split = "")[[1]], dehex)
  names(L) <- paste("Q", 1:questions, sep=".")
	L
  }
}

# need to deal with NA right above and below
test <- function (x) {
	questions <- max(sapply(as.character(x), nchar))
	data.frame(t(sapply(x, function(y) grades(y, questions))))
}

makeTest <- function (q) {
  function(x) test(x, q)
}
```
```{r}
# data ingestion
# START ROW DEPENDS ON CATALYST
# END ROW DEPENDS ON NUMBER OF STUDENTS
# WEBASSIGN EXPORT IS ESSENTIALLY BROKEN, SO HAVE TO CONVERT TO CSV
# THEN THOSE MORONS ENCODED A COLUMN OF PERCENTAGES USING %!!!!
cats <- read.xlsx("catalyst3.xls", 1, colClasses="character", startRow=3)[-c(1,2),]
cats <- cats[cats$Status == "Active",]
hwk <- read.xlsx("homework.xls", 1, colClasses="character", rowIndex=9:184, colIndex=1:4)
homeworktotal <- 618.25 # HORRIBLE MANUAL CHECK AND SET
hwk$Homework <- as.numeric(hwk[,4]) / 618.25
#hwk$Username <- sub("@.*", "", hwk$Username)
all <- merge(cats, hwk, by="Username")
all$Homework[is.na(all$Homework)] <- 0
```

Here's a summary of how the class did on all of the exams.

## Midterm 1

```{r}
mid.1 <- test(all$Midterm.1)
qs <- ncol(mid.1)
mid.1$total <- rowSums(mid.1)
curr.test <- mid.1
mid.1.points <- 40
# exclude the 0s; this probably means something was very wrong, like students absent
curr.test <- curr.test[curr.test$total != 0 & complete.cases(curr.test),]
```

There were a total of 40 points.

The median is `r median(curr.test$total)`, the mean is `r mean(curr.test$total)`, and the standard deviation is `r sd(curr.test$total)`. Here are a few plots showing distributions:

```{r}
ggplot(melt(curr.test[,1:qs])) + geom_boxplot(aes(x=variable, y=value)) + xlab("Problem") + ylab("Points")
```
```{r}
ggplot(melt(curr.test[,1:qs])) + geom_histogram(aes(x=value), binwidth=1) + facet_grid(.~variable)
```

```{r}
ggplot(curr.test) + geom_histogram(aes(x=total), binwidth=2)
```

This last plot compares the probability density derived from class performance with a normal distribution having the same mean and standard deviation; it is a smoother representation of how the class distribution compares to a normal.

```{r}
df <- curr.test[curr.test$total != 0, ]
ggplot(df, aes(x=total)) + geom_density() + 
  stat_function(fun=dnorm, args=list(mean=mean(df$total), sd=sd(df$total)), color='red')
```


## Midterm 2

```{r}
mid.2 <- test(all$Midterm.2)
qs <- ncol(mid.2)
mid.2$total <- rowSums(mid.2)
curr.test <- mid.2
mid.2.points <- 40
# exclude the 0s; this probably means something was very wrong, like students absent
curr.test <- curr.test[curr.test$total != 0  & complete.cases(curr.test),]
```
There was a total of 40 points.

The median is `r median(curr.test$total)`, the mean is `r mean(curr.test$total)`, and the standard deviation is `r sd(curr.test$total)`. Here are a few plots showing distributions:

```{r}
ggplot(melt(curr.test[,1:qs])) + geom_boxplot(aes(x=variable, y=value)) + xlab("Problem") + ylab("Points")
```
```{r}
ggplot(melt(curr.test[,1:qs])) + geom_histogram(aes(x=value), binwidth=1) + facet_grid(.~variable)
```

```{r}
ggplot(curr.test) + geom_histogram(aes(x=total), binwidth=2)
```

This last plot compares the probability density derived from class performance with a normal distribution having the same mean and standard deviation; it is a smoother representation of how the class distribution compares to a normal.


```{r}
df <- curr.test[curr.test$total != 0, ]
ggplot(df, aes(x=total)) + geom_density() + 
  stat_function(fun=dnorm, args=list(mean=mean(df$total), sd=sd(df$total)), color='red')
```


## Final exam

```{r}
final <- test(all$Final)
qs <- ncol(final)
final$total <- rowSums(final)
curr.test <- final
final.points <- 100
# exclude the 0s; this probably means something was very wrong, like students absent
curr.test <- curr.test[curr.test$total != 0  & complete.cases(curr.test),]
```

There were a total of 100 points.

The median is `r median(curr.test$total, na.rm=TRUE)`, the mean is `r mean(curr.test$total, na.rm=TRUE)`, and the standard deviation is `r sd(curr.test$total, na.rm=TRUE)`. Here are a few plots showing distributions:

```{r}
ggplot(melt(curr.test[,1:qs])) + geom_boxplot(aes(x=variable, y=value)) + xlab("Problem") + ylab("Points")
```
```{r}
ggplot(melt(curr.test[,1:qs])) + geom_histogram(aes(x=value), binwidth=2) + facet_grid(.~variable)
```

```{r}
ggplot(curr.test[curr.test$total != 0, ]) + geom_histogram(aes(x=total), binwidth=4)
```

This last plot compares the probability density derived from class performance with a normal distribution having the same mean and standard deviation; it is a smoother representation of how the class distribution compares to a normal.


```{r}
df <- curr.test[curr.test$total != 0, ]
ggplot(df, aes(x=total)) + geom_density() + 
  stat_function(fun=dnorm, args=list(mean=mean(df$total), sd=sd(df$total)), color='red')
```


## Aggregated test scores

```{r}
valid.students <- (final$total != 0 & complete.cases(final))
m1 <- mid.1$total * 27.5 / mid.1.points
m2 <- mid.2$total * 27.5 / mid.2.points
m3 <- final$total * 45.0 / final.points

m1.0 <- m1 == 0
m2.0 <- m2 == 0

multiplier <- 4/2 * (2 - m1.0 - m2.0)
multiplier[multiplier == 2] <- 1

test.grade <- m1 + m2 + m3
test.grade <- test.grade * multiplier
test.grade.valid <- test.grade[valid.students]
```

Here are the aggregate the data for all exams (given equal weight), normalized to be out of 100. The median is `r median(test.grade.valid)`, the mean is `r mean(test.grade.valid)`, and the standard deviation is `r sd(test.grade.valid)`.

Here is a histogram showing the curve resulting from aggregating all tests (with equal weight), normalized to be out of 100. 

```{r}
df <- data.frame(total = test.grade.valid)
ggplot(df, aes(x=total)) + geom_histogram()
```

This last plot compares the probability density derived from class performance with a normal distribution having the same mean and standard deviation; it is a smoother representation of how the class distribution compares to a normal. You can see that the right-skewed midterms overcome the almost-normal final exam to give a slightly right-skewed aggregate.

```{r}
df <- data.frame(total = test.grade.valid)
ggplot(df, aes(x=total)) + geom_density() + 
  stat_function(fun=dnorm, args=list(mean=mean(test.grade.valid), sd=sd(test.grade.valid)), color='red')
```


## Homework
```{r}
all$Homework.adjustment <- (all$Homework >= 0.895) * 0.1 - (all$Homework <= 0.600) * 0.1
all.valid <- all[valid.students,]
```
Homework adjustments for the final grade were either 0.1 or -0.1, depending upon whether a student got at at least 89.5% of the homework points or at most 60% of the homework points. 
The distribution of homework grades (excluding students who got 0 on the final), with lines marking the adjustment boundaries, looks like this.
```{r}
ggplot(all.valid) + geom_histogram(aes(x=Homework), binwidth=0.04) + geom_vline(xintercept=0.60, color="red") + geom_vline(xintercept=0.895, color="green")
```
Among students who took the final, `r sum(all.valid$Homework.adjustment > 0)` students received a positive adjustment, while `r sum(all.valid$Homework.adjustment < 0)` students received the negative adjustment.

<!--
## Journal
```{r}
#all$Journals <- as.numeric(as.character(all$Journals))
#all.valid <- all[valid.students,]
```
Journal adjustments where 0.2, 0.1, 0.0, -0.1, -0.2, depending upon the quality of the journal entries. The distribution of adjustments looks like this.

```{r}
#ggplot(all.valid) + geom_histogram(aes(x=Journals), binwidth=0.1) + xlim(-0.2, 0.3)
```
In particular, among students who took the final, `r sum(all.valid$Journals < 0)` got a negative adjustment and `r sum(all.valid$Journals > 0)` got a positive adjustment.
-->

## The final curve
```{r}
raw.test.curve <- ((test.grade - mean(test.grade.valid, na.rm=TRUE)) / (2 * sd(test.grade.valid, na.rm=TRUE)) + 2.9)
adjustments <- all$Homework.adjustment #+ all$Journals 
final.curve <- raw.test.curve + adjustments
final.curve[final.curve > 4.0] <- 4.0
final.curve[final.curve < 0.0] <- 0.0
final.curve[!valid.students] <- 0.0
all$final.grade <- round_any(final.curve, 0.1)
write.csv(all, "final_grades.csv")
```

Here's the final class curve, with the median dashed and the mean dotted. For comparison, a normal distribution with the same mean and median is also shown in red. Look at those bulbous tails!

```{r}
df <- data.frame(curve=final.curve[valid.students])
ggplot(df, aes(x=curve)) + geom_density() + geom_vline(xintercept=median(df$curve), linetype="dashed") + geom_vline(xintercept=mean(df$curve), linetype="dotted") + stat_function(fun=dnorm, args=list(mean=mean(df$curve), sd=sd(df$curve)), color='red')
```